---
title: "Analysis: framing demo"
author: "Your name"   # <- put your name here
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
library(tidyverse)
library(jsonlite)
```

This notebook reads the CSVs written by `npm run export -- --experiment framing-demo`
and runs the planned analyses for the demo experiment. Set `experiment_id` in the first
code chunk below to match `EXPERIMENT.id` in `experiment.js`, and put your name in the
`author:` line at the top of this file. Replace the sections below with
the confirmatory analysis you preregistered; keep exploratory analyses clearly separated.

## Load data

`trials.csv` is long format: one row per jsPsych trial, one participant per `participant_id`.
Columns that hold objects (for example the survey plugin's `response`) are JSON strings.

```{r load}
experiment_id <- "framing-demo"
raw_dir <- file.path("..", "data", "raw", experiment_id)

participants <- read_csv(file.path(raw_dir, "participants.csv"), show_col_types = FALSE)
trials <- read_csv(file.path(raw_dir, "trials.csv"), show_col_types = FALSE)
errors <- read_csv(file.path(raw_dir, "errors.csv"), show_col_types = FALSE)

glimpse(participants)
```

## Exclusions

Document every exclusion. Record how many participants each rule removes.

```{r exclusions}
n_started <- nrow(participants)
complete <- participants |> filter(completed)

# Example attention/performance criterion: at least 6 of 8 lexical decision trials correct.
# Replace this with the criterion you preregistered.
ld_accuracy <- trials |>
  filter(task == "lexical_decision") |>
  group_by(participant_id) |>
  summarise(
    ld_acc = mean(as.logical(correct), na.rm = TRUE),
    ld_all_timed_out = all(as.logical(timed_out)),
    .groups = "drop"
  )

# A participant who never answered a single speeded trial did not really take part: this
# is what a phone or tablet produces, and what an abandoned-but-left-open tab produces.
keep <- complete |>
  left_join(ld_accuracy, by = "participant_id") |>
  filter(!is.na(ld_acc), !ld_all_timed_out, ld_acc >= 0.75)

# Funnel by condition. `.drop = FALSE` on a factor keeps zero rows visible: a step that
# removes everyone must show as 0, not vanish from the table.
funnel_levels <- c("started", "completed", "passed accuracy criterion")
bind_rows(
  participants |> mutate(step = "started"),
  complete |> mutate(step = "completed"),
  keep |> mutate(step = "passed accuracy criterion")
) |>
  mutate(step = factor(step, levels = funnel_levels)) |>
  count(step, condition, .drop = FALSE) |>
  pivot_wider(names_from = condition, values_from = n, values_fill = 0) |>
  arrange(step)

message("Participants: ", n_started, " started / ", nrow(complete), " completed / ",
        nrow(keep), " kept for analysis.")

# Participants excluded because they never responded to a speeded trial (e.g. no keyboard).
n_no_response <- complete |> left_join(ld_accuracy, by = "participant_id") |>
  filter(ld_all_timed_out %in% TRUE) |> nrow()
if (n_no_response > 0) {
  message("Note: ", n_no_response,
          " completed session(s) had no response on any speeded trial. Check whether they ",
          "were taken on a device without a keyboard.")
}

if (nrow(keep) == 0) {
  message("No participants passed the exclusion criteria, so the analyses below will be ",
          "empty. With pilot data this usually just means too few participants so far.")
}
```

## Demographics

```{r demographics}
demo <- trials |>
  filter(task == "demographics", participant_id %in% keep$participant_id) |>
  mutate(resp = map(response, ~ fromJSON(.x))) |>
  transmute(
    participant_id,
    age = map_dbl(resp, ~ as.numeric(.x$age %||% NA)),
    gender = map_chr(resp, ~ as.character(.x$gender %||% NA)),
    native_english = map_lgl(resp, ~ isTRUE(.x$native_english))
  )

demo |> summarise(n = n(), mean_age = mean(age, na.rm = TRUE), sd_age = sd(age, na.rm = TRUE))
demo |> count(gender)
```

## Confirmatory analysis: framing effect

Key test (preregistered): proportion choosing the risky program (B) is higher in the loss
frame than in the gain frame. Chi-square test of independence on the 2 x 2 table.

```{r framing}
framing <- trials |>
  filter(task == "framing", participant_id %in% keep$participant_id) |>
  select(participant_id, condition, choice)

framing_counts <- framing |> count(condition, choice) |>
  group_by(condition) |> mutate(prop = n / sum(n)) |> ungroup()
framing_counts

tab <- table(framing$condition, framing$choice)
tab

if (all(dim(tab) == c(2, 2)) && all(rowSums(tab) > 0)) {
  print(chisq.test(tab))
} else {
  message("Chi-square not run: the frame-by-choice table is ",
          paste(dim(tab), collapse = " x "),
          ". It needs both frames and both choices to be present. ",
          "With ", nrow(framing), " participant(s) in the analysis set this is expected; ",
          "come back once you have more data.")
}

# Effect size with a confidence interval: the difference in the proportion choosing the
# risky program (loss minus gain). Compare this interval with the original effect, not
# just the p value.
risky <- framing |>
  group_by(condition) |>
  summarise(risky = sum(choice == "risky"), n = n(), .groups = "drop")
risky

have_both <- all(c("gain", "loss") %in% risky$condition)
if (have_both && all(risky$n > 0)) {
  # Index each frame explicitly. Do NOT write risky$risky[risky$condition == c("loss","gain")]:
  # that compares element by element, silently returns nothing, and prop.test then fails.
  loss <- risky[risky$condition == "loss", ]
  gain <- risky[risky$condition == "gain", ]
  pt <- prop.test(x = c(loss$risky, gain$risky), n = c(loss$n, gain$n), correct = FALSE)
  print(pt)
  tibble(
    prop_risky_loss = loss$risky / loss$n,
    prop_risky_gain = gain$risky / gain$n,
    difference_loss_minus_gain = unname(pt$estimate[1] - pt$estimate[2]),
    ci_low = pt$conf.int[1],
    ci_high = pt$conf.int[2]
  )
} else {
  message("Proportion difference not computed: need at least one kept participant in each ",
          "frame (have ", paste(paste0(risky$condition, "=", risky$n), collapse = ", "), ").")
}

if (nrow(framing_counts) > 0) {
  ggplot(framing_counts, aes(x = condition, y = prop, fill = choice)) +
    geom_col() +
    labs(x = "Frame", y = "Proportion of choices", fill = "Choice",
         title = paste0("Framing effect (n = ", nrow(framing), ")")) +
    theme_minimal()
} else {
  message("No framing data to plot yet.")
}
```

## Exploratory: lexical decision RTs

```{r rt}
ld <- trials |>
  filter(task == "lexical_decision", participant_id %in% keep$participant_id) |>
  mutate(correct = as.logical(correct), is_word = as.logical(is_word), rt = as.numeric(rt))

ld_correct <- ld |> filter(correct, !is.na(rt))

if (nrow(ld_correct) > 0) {
  print(ld_correct |> group_by(is_word) |>
          summarise(mean_rt = mean(rt), sd_rt = sd(rt), n = n(), .groups = "drop"))
  ggplot(ld_correct, aes(x = rt, fill = is_word)) +
    geom_histogram(bins = 30, alpha = 0.6, position = "identity") +
    labs(x = "RT (ms)", fill = "Real word") +
    theme_minimal()
} else {
  message("No correct lexical-decision trials with a response time yet.")
}
```

## Data quality

Client-side errors reported by the experiment page, if any.

```{r errors}
if (nrow(errors) > 0) {
  errors |> select(any_of(c("at", "uid", "trial_index", "message"))) |> head(20)
} else {
  message("No client-side errors were reported. This is what you want to see.")
}
```

## Session info

```{r session}
sessionInfo()
```
